---
title: "Data Analysis Memo"
author: "Hugo Novales"
date: "2026-04-18"
output:
  pdf_document:
    extra_dependencies: ["float"]
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.pos = "H")

library(ggplot2)
library(tidyverse)
library(kableExtra)
library(plm)
library(lmtest)
library(sandwich)
library(sjPlot)
library(sensemakr)
library(readxl)
library(vdemdata)
library(modelsummary)
library(huxtable)
library(lmtest)
library(sandwich)

```

# 1. Research Question

What explains variation in majority formation in multiparty legislatures? There is consensus in political science about the role of district magnitude and party fragmentation (Cox 1997; Lijphart 1999; Norris 2004). However, does this rationale extend to majority formation? This question is important because it speaks to the actual performance of legislatures. It is also important because majority formation is how political and policy change occurs in democracies. If majorities are not formed in legislatures, then democracy may stagnate and fail to perform in a manner that is satisfactory for citizens.
District magnitude is relatively stable, as electoral rules are a matter of contention. Comparing across cases is difficult because specific countries may have different drivers of majority formation or gridlock.

The causal logic is important. Electoral systems do not in and of themselves produce more parties, but, where multiple parties exist, they may repress the number of parties or fail to do so. District magnitude here is a key variable. Cox (1997)  and Sartori before him have found that smaller district magnitudes tend to exert pressure on the number of parties, while larger magnitudes relax this pressure and allow for more proliferation.

The number of parties and their relative sizes (fragmentation, effective number of parties) may be associated to majority formation (Mainwaring 1993; Tsebelis 1995). However, fragmentation does not say anything about whether those parties have substantial reasons to agree on policy. Polarization, therefore, could also influence majority formation. While I'm not interested in the effect of polarization, it is important to consider it in a causal model.

I have chosen to focus my analysis on a single case. The first reason is that this allows me to control for noise associated to national context variation. In other words, there may be a reason why politicians in country X tend to form majorities less frequently than politicians in neighboring country Y. It could be ethnic cleavages, political culture, legislative procedural rules, or other unknown factors. Focusing on a single country allows me to isolate my variables of interest, at least to a larger measure than if I were employing a cross-country analysis.

While we know that restrictive electoral systems  -such as those with smaller district magnitude-  tend to repress or suppress the number of parties, it is rare that a country reduces its district magnitude. Legislators are typically in charge of electoral reform, and they reasonably understand that reducing district magnitude may act against their probability of reelection (Andrews and Jackman 2005; Negretto and Visconti 2018). Therefore, overwhelmingly, district magnitude tends to increase rather than decrease. This presents a problem for identification, as the effects of electoral system reform may linger into the future. Societal fragmentation or political polarization  -or any other idiosyncratic variable that affects majority formation- may increase or decrease organically without changes in the electoral system.
A better test for the effects of electoral system change on majority formation is when changes in opposite directions occur close in time.  Guatemala is one such case. Since the first democratic election in the current regime in 1986, over the course of ten elections and an equal number of legislatures, district magnitude has changed 5 times, increasing 4 times and decreasing once (Brolo 2012; Brolo 2013). This within-case variation allows me to approximate a causal effect much more than in a country where magnitude has only increased. Furthermore, restricting the analysis to a single case with such a short democratic spell sets a high bar for causality: since the number of observations is very small, each single observation should be understood to have enough leverage so that changes in the dependent variable attributable to factors other than my explanation should significantly weaken my theory.


# 2.  Causal Model and Hypotheses

My theory is as follows. District magnitude is the single most important
variable that influences majority formation. First, it will increase the
number of parties -presumably with different views- present in a
legislature. Second, in increasing the number of parties and creating a
more proportional seat allocation, it will make the legislature a more
faithful reflection of polarization in society, making majority
formation more difficult the more polarized society is. That is, the
effect of district magnitude is sensitive to polarization: when
polarization increases, the effect should be stronger. Finally, I expect
majority formation to be sensitive to political events. One specific
type of political event is elections. In the year following an election,
when the legislature is newly integrated, they may be motivated to
deliver on campaign promises and party cohesion is expected to be
strongest. As the next election year approaches, legislators will be
more occupied with campaigning, therefore affecting legislative output.
This means that my causal model should control for distance to the prior
or next election.

Hypothesis 1

Higher district magnitude is associated with lower frequency of majority
formation.

Hypothesis 2

The effect of district magnitude is partially mediated by party
fragmentation.

Hypothesis 3

District magnitude interacts with societal polarization to decrease the
frequency of majority formation.

Hypothesis 4

Majority formation will be less frequent as the next election
approaches.

I have modeled these hypothesis as follows:

Majority formation = district magnitude + party fragmentation +
polarization + distance to next election + (polarization \* district
magnitude)

# 3.  Data

The effect of district magnitude on majority formation would be best
tested in a controlled environment by setting up n polities and
assigning them different district magnitudes, observing them over a
constant period of time and counting the frequency with which voting
majorities form in the legislature. This is not possible. Instead, I
have chosen a single polity -Guatemala-, which has some variation in
district magnitude over time. This is the main independent variable. One
way to measure the dependent variable -majority formation- would be to
analyze voting records to identify the frequency with which each
legislature achieves a majority. I don't have access to voting records,
but I was able to scrape the website of the Guatemalan legislature for
every piece of legislation approved since 1928. During my period of interest, each
piece of legislation is officialy coded with a decree number ("número de
decreto") which includes a index number ranging from 1 to the number of
laws passed in a year, followed by a hyphen and the year in which the
legislation was passed. This allows me to count the number of pieces of
legislation approved in each year. I am then using legislative ouput as
a proxi for majority formation.

Guateamala experienced regime change several times over the course of the
20th century. However, after the democratic election in 1985, the regime
has been reasonably stable, holding legislative elections which are
reasonably free and fair every five years until 1995, and every four
years thereafter. This allows me to use seat allocation data from each
of those elections to build a party fragmentation measure. I have
obtained this data from the V-Party Dataset by the V-Dem project.
I have also obtained data on political polarization from the V-Dem
database.

```{r, data cleaning, organizing, wrangling, include=FALSE}

#I downloaded the source code from https://www.congreso.gob.gt/seccion_informacion_legislativa/decretos and stored it as a text file in source_code_decretos_congreso

decretos<-paste(readLines("source_code_decretos_congreso"), collapse = "\n")
# capture "de 1994", "del 2023", etc.
pattern <- "\\bde[l]?\\s+(19|20)\\d{2}\\b"

phrases <- str_extract_all(decretos, pattern)[[1]]

# extract the year out of each phrase
years <- str_extract(phrases, "(19|20)\\d{2}")

decrees_per_year <- data.frame(year = years) %>%
  count(year, name = "n") %>%
  arrange(as.numeric(year))

decrees_per_year_democracy<-decrees_per_year%>%filter(year>=1986)
decrees_per_year_democracy$ELECTION<-c(1985, 1985, 1985, 1985, 1985, 
                                       1990, 1990, 1990, 1990, 1990, 
                                       1995, 1995, 1995, 1995, 
                                       1999, 1999, 1999, 1999, 
                                       2003, 2003, 2003, 2003, 
                                       2007, 2007, 2007, 2007, 
                                       2011, 2011, 2011, 2011, 
                                       2015, 2015, 2015, 2015, 
                                       2019, 2019, 2019, 2019, 
                                       2023, 2023)

decrees_per_year_democracy<-as.tibble(decrees_per_year_democracy)
decrees_per_year_democracy$ELECTION<-as.double(decrees_per_year_democracy$ELECTION)
decrees_per_year_democracy
decrees_per_year

district_magnitude <- data.frame(
  District = c(
    "El Progreso","Baja Verapaz","Zacapa","Sacatepequez","Santa Rosa","Sololá",
    "Retalhuleu","Izabal","Chiquimula","Jalapa","Totonicapán","Petén","Jutiapa",
    "Chimaltenango","Suchitepequez","Escuintla","Quetzaltenango","Quiché",
    "San Marcos","Alta Verapaz","Huehuetenango","Central","Guatemala",
    "National List"
  ),
  `1985` = c(2,2,2,2,2,2,2,2,2,2,2,2,3,2,2,4,5,4,6,4,5,10,6,25),
  `1990` = c(2,2,2,2,3,2,2,3,3,2,3,2,4,3,3,5,5,5,6,5,6,10,7,29),
  `1995` = c(1,1,1,1,2,1,1,2,2,1,2,1,3,2,2,4,4,4,5,4,5,9,6,16),
  `1999` = c(1,1,1,2,3,2,2,3,2,2,3,2,3,3,3,4,6,5,8,6,7,10,12,22),
  `2003` = c(1,2,2,3,3,3,3,3,3,3,4,4,4,5,5,6,7,8,9,9,10,11,19,31),
  `2007` = c(1,2,2,3,3,3,3,3,3,3,4,4,4,5,5,6,7,8,9,9,10,11,19,31),
  `2011` = c(1,2,2,3,3,3,3,3,3,3,4,4,4,5,5,6,7,8,9,9,10,11,19,31),
  `2015` = c(1,2,2,3,3,3,3,3,3,3,4,4,4,5,5,6,7,8,9,9,10,11,19,31),
  `2019` = c(2,2,2,3,3,3,3,3,3,3,4,4,4,5,5,6,7,8,9,9,10,11,19,32),
  `2023` = c(2,2,2,3,3,3,3,3,3,3,4,4,4,5,5,6,7,8,9,9,10,11,19,32),
  check.names = FALSE
  )

district_magnitude

avg_magnitude<- district_magnitude %>%
  pivot_longer(-District, names_to = "ELECTION", values_to = "MAGNITUDE") %>%
  mutate(ELECTION = as.double(ELECTION)) %>%
  group_by(ELECTION) %>%
  summarize(avg_magnitude = mean(MAGNITUDE))

avg_magnitude


seats_per_party_per_election <- read_excel("seats_per_party_per_election.xlsx")
seats_per_party_per_election

ENP <- seats_per_party_per_election %>%
  group_by(ELECTION) %>%
  summarize(
    ENP = 1 / sum(SEAT_SHARE^2)
  )

ENP

vparty_guate_democracy<-vparty%>%
  filter(country_name == "Guatemala")%>%
  filter(year >= 1985)
vparty_guate_democracy
guate_democracy_seats<-vparty_guate_democracy%>%
  select(v2paenname, v2pashname, year, v2panumbseat)
  
guate_democracy_seats


summary(decrees_per_year)
ggplot(data = decrees_per_year_democracy, aes(x = year, y = n)) + geom_col()
decrees_per_magnitude<-left_join(decrees_per_year_democracy, avg_magnitude, by = "ELECTION")
decrees_per_magnitude <- decrees_per_magnitude %>%
  group_by(ELECTION) %>%
  mutate(year_in_term = row_number()) %>%
  ungroup()
decrees_per_magnitude
polarization_gt <- vdem %>%
  filter(country_name == "Guatemala") %>%
  select(year, v2cacamps_osp)%>%filter(year >= "1986")
polarization_gt
polarization_gt$year<-as.character(polarization_gt$year)
decrees_per_magnitude<-left_join(decrees_per_magnitude, polarization_gt, by = "year")
decrees_per_magnitude<-left_join(decrees_per_magnitude, ENP, by = "ELECTION")
summary(decrees_per_magnitude)

```

```{r, descriptive exploration, table 1, echo = FALSE, message = FALSE, warning = FALSE}

kable(district_magnitude, caption = "Districts and District Magnitude in Guatemala, 1985-2025")

```

\newpage

```{r, Table 2, echo = FALSE, message = FALSE, warning = FALSE}
kable(avg_magnitude, caption = "Average District Magnitude by Election in Guatemala, 1985-2025")

```

\newpage

```{r, Table 3, echo = FALSE, message = FALSE, warning = FALSE}
kable(decrees_per_year_democracy, caption = "Legislative Output (n) by Year in Guateamla, 1986-2025")

```

\newpage

```{r, figure 1, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Average District Magnitude by Election"}
ggplot(avg_magnitude, aes(ELECTION, avg_magnitude))+ geom_col() 

```

\newpage

```{r, figure 2, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Effective Number of Parties by Election"}

ggplot(decrees_per_magnitude, aes(x = factor(ELECTION), y = ENP)) + geom_col(stat = "summary", fun = "mean")

```

\newpage

```{r figure 3, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Polarization (v2cacamps_ops) by Year (V-Dem)"}

ggplot(decrees_per_magnitude, aes(year, v2cacamps_osp, group = 1))+ 
  geom_line() + 
  theme(axis.text.x = element_text(angle = 90))
```

\newpage

```{r, figure 4, fig.cap = "Number of Decrees (n) by Year"}
ggplot(data = decrees_per_magnitude, aes(x = year, y = n)) + 
  geom_col() + 
  facet_wrap(~ELECTION, scales = "free_x")

```

```{r, figure 5, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Number of Decrees by Effective Number of Parties" }

ggplot(data = decrees_per_magnitude, aes (x = ENP, y = n)) + geom_point()
```

```{r, figure 6, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Number of Decrees (n) by Average District Magnitude"}

ggplot(data = decrees_per_magnitude, aes (x = avg_magnitude, y = n)) + geom_point()

```

```{r, figure 7, echo = FALSE, message = FALSE, warning = FALSE, fig.cap = "Number of Decrees (n) by Year in Term"}

ggplot(data = decrees_per_magnitude, aes (x = year_in_term, y = n)) + geom_point()

```


\newpage

# 4.  Analysis

I ran an OLS regression for my main independent variable -average district magnitude- and my main dependent variable -legislative output- (Model 1), and then I control for year in term and polarization in subsequent models (2 and 3. Based on my model  model described above, I included an interaction term where political polarizaton (v2cacamps_obs) interacts with district magnitude (Model 4).  I find that year in term explains a small part of legislative output across all models. In models 1-3, district magnitude is the main driver behind legislative output in the anticipated direction. However, in the last model, the interaction term moves all this explanatory weight to polarization, and significance dissappears. Model 2 offers the most parsimonious explanation and is therefore the one I chose to continue working with below. Thus, I find support for Hypothesis 1 and Hypothesis 4, but not for Hypothesis 3. 

```{r, analysis, echo = FALSE, message = FALSE, warning = FALSE, results = 'hide'}
model1<-lm(n~avg_magnitude, decrees_per_magnitude)
summary(model1)
coeftest(model1)
coeftest(model1, vcov = vcovCL, cluster = ~ELECTION)

model2<-lm(n~avg_magnitude + year_in_term, decrees_per_magnitude)
model2
summary(model2)
coeftest(model2)
coeftest(model2, vcov = vcovCL, cluster = ~ELECTION)

model3<-lm(n~avg_magnitude + year_in_term + v2cacamps_osp, decrees_per_magnitude)
model3
summary(model3)
coeftest(model3)
coeftest(model3, vcov = vcovCL, cluster = ~ELECTION) #it's not polarization. wohooo. 

model4<-lm(n~avg_magnitude + year_in_term + v2cacamps_osp + v2cacamps_osp*avg_magnitude, decrees_per_magnitude)
coeftest(model4)
coeftest(model4, vcov = vcovCL, cluster = ~ELECTION)
#significantly deflates avg_magnitude and eliminates significance

```

```{r, echo=FALSE}
modelsummary(list ("Model 1" = model1, "Model 2" = model2, "Model 3" = model3, "Model 4" = model4), stars = TRUE, title = "Results of OLS Regression for Specified Model")

```

```{r, echo = FALSE, fig.cap="Legislative Output Regressed on Average District Magnitude"}
plot_model(model2, type = "eff", terms = "avg_magnitude") +
  theme_minimal()


```

I then proceed to integrate the Effective Number of Parties (ENP) into the model. As we can see from the results below, the inclusion of ENP significantly reduces the weight of district magnitude. District magnitude is highly correlated with ENP, and we know from conventional literature that district magnitude impacts party fragmentation. Therefore, ENP is assumed here to be a mediating variable and is thus discarded from the model. Thus, I find support for hypothesis 2. 


```{r, with ENP, echo = FALSE, message = FALSE, results = 'hide', warning = FALSE}


model1ENP<-lm(n~avg_magnitude + ENP, decrees_per_magnitude)

model2ENP<-lm(n~avg_magnitude + year_in_term + ENP, decrees_per_magnitude)


model3ENP<-lm(n~avg_magnitude + year_in_term + v2cacamps_osp + ENP, decrees_per_magnitude)

```

```{r, echo = FALSE}

modelsummary(list ("Model 1" = model1ENP, "Model 2" = model2ENP, "Model 3" = model3ENP), stars = TRUE, title = "Results of OLS Regression for Specified Model Controlling for Effective Number of Parties")

modelX<-lm(ENP~avg_magnitude, decrees_per_magnitude)

modelsummary(modelX, stars = TRUE, title = "Effective Number of Parties Regressed on District Magnitude")


```

\newpage

# 5.  Assumption Review and Robustness Checks

The following regression diagnostics plots allow me to review the necessary assumptions for my OLS identification strategy. I am using Model 2. The Residuals vs. Fitted plot shows a reasonably stable line and identifies certain outliers (22, 20, 14). The line is slightly curved (U-Shaped) suggesting non-linnearity. Similarly, in the  Q-Q Residuals plot, while all points lie along the diagonal, the same points as in the Residuals vs. Fitted test lie outside of the diagonal, suggesting that residuals are may not be normally distributed. The Scale-Location plot shows reasonable uniformity in the distribution of all points, suggesting homoscedasticity. Finally, in the Residuals vs. Leverage plot, while   the aforementioned outliers have more leverage than the rest of the points, they do not cross into the significance corners. 


```{r, tests and assumptions, plot, echo = FALSE, message = FALSE, results = 'hide', warning = FALSE}
##### Abajo los plots para hacer los tests


#par(mfrow = c(2, 2)) #
plot(model2)
#par(mfrow = c(1, 1)) # 

```
Since the plots above suggest possible non-linearity and non-normality, I ran statistical tests to check for each of the OLS assumptions. The results are described in Table 7. Overall, Model 2 passes all tests except for the Durbin-Watson test. Observations are not independent, mainly because one important driver of change is the integration of a new legislature with every election. Different legislatures will have different ideological configurations; in some legislatures, the president's party has a majority while in others it does not; also, different issues may come up at different times over 40 years and affect polarization, policy choices and the disposition to form majorities. 

Table 8 shows the results for the same models that I specified above, adjusting standard errors to account for clustering around the variable "ELECTION", which defines each legislative session. Coefficients and significance levels remain similar. 

```{r, assumption checks, echo = FALSE, message = FALSE, results = 'hide', warning = FALSE, include = FALSE}

Linearity<-raintest(model2) # p-value > 0.05 Reject Linearity H0 (passed test)
Linearity

IID<-dwtest(model2) #p-value < 0.05. True autocorrelation is greater than 0, meaning dependencia temporal (does not pass IID sample test)
IID

Multicollinearity<-car::vif(model2) #no hay multicolinearidad, o al menos no hay demasiada...? entre dm y enp puede haber, pero no es total. pasa
Multicollinearity

Homoscedasticity<-bptest(model2) # p-value = 0.7683, H0, no hay heteroscedasticidad (hay homocedasticidad), pasa. 
Homoscedasticity

Normality<-shapiro.test(residuals(model2)) #p-value 0.246, H0, pasa
Normality









#abajo de aquí estoy guardando por cualquier cosa, pero creo que no sirve. 

#panel_decrees_per_magnitude<-pdata.frame(decrees_per_magnitude, index = c("ELECTION", "year"))
#panel_decrees_per_magnitude
 # model3<-plm(n~avg_magnitude, decrees_per_magnitude,index = c("ELECTION", "year"), 
#            model = "random", 
 #           random.method = "nerlove") #This does not work, with any random.method. Explanation: no within-ELECTION variation. 

# causal workflow
# Question
# Estimand
# Ideal experiment 
# Identification strategy
# Estimation
# Inference / uncertainty


```

**Table 7: Assumption Review**

| Assumption | Test | Statistic | P-Value | Result |
|---|---|---|---|---|
| Linearity | Rainbow | F = 0.563 | 0.891 | Pass |
| Independence | Durbin-Watson | DW = 1.526 | 0.048 | Fail |
| Homoscedasticity | Breusch-Pagan | BP = 0.387 | 0.824 | Pass |
| Normality | Shapiro-Wilk | W = 0.970 | 0.367 | Pass |
| Multicollinearity | VIF | avg_magnitude = 1.016, year_in_term = 1.016 | — | Pass |

```{r, coefficient tests, echo = FALSE, message = FALSE, warning = FALSE, message = FALSE, results = 'hide', warning = FALSE}

coeftest1<- coeftest(model1, vcov = vcovCL, cluster = ~ELECTION)
coeftest2<- coeftest(model2, vcov = vcovCL, cluster = ~ELECTION)
coeftest3<- coeftest(model3, vcov = vcovCL, cluster = ~ELECTION)
coeftest4<- coeftest(model4, vcov = vcovCL, cluster = ~ELECTION)

```

```{r coeftest table}

modelsummary(list ("Model 1" = coeftest1, "Model 2" = coeftest2, "Model 3" = coeftest3, "Model 4" = coeftest4), stars = TRUE, title = "Coefficient Tests for Specified Models on ELECTION Cluster")

```


\newpage

# 7.  Bibliogrpahy

Brolo, Javier. 2012. Resultados electorales de los partidos políticos en Guatemala de 1985 a 2012. Guatemala City: ASIES.

Brolo, Javier. 2013. El tamaño importa: Variaciones por distrito en elecciones legislativas de Guatemala (1984–2011). Guatemala City: ASIES.

Carey, John M., and Simon Hix. 2011. "The Electoral Sweet Spot: Low-Magnitude Proportional Electoral Systems." American Journal of Political Science 55(2): 383–97.


Cox, Gary W. 1997. Making Votes Count: Strategic Coordination in the World's Electoral Systems. Cambridge: Cambridge University Press.

Figueroa, Valentin. 2025. "Does Proportionality Increase Turnout? A Study of Adaptation to Oscillating Electoral Systems." Comparative Political Studies OnlineFirst. https://doi.org/10.1177/00104140241290209.

Mainwaring, Scott. 1993. "Presidentialism, Multipartism, and Democracy: The Difficult Combination." Comparative Political Studies 26(2): 198–228.

Negretto, Gabriel L., and Giancarlo Visconti. 2018. "Electoral Reform under Limited Party Competition: The Adoption of Proportional Representation in Latin America." Latin American Politics and Society 60(1): 27–51.

Naslund, Karlerik. 2025. "Regression and Testing Cheat Sheet." RPubs. May 7. https://rpubs.com/knaslundjr/cheat_sheet_2.

Norris, Pippa. 2004. Electoral Engineering: Voting Rules and Political Behavior. Cambridge: Cambridge University Press.

Tsebelis, George. 1995. "Decisionmaking in Political Systems: Veto Players in Presidentialism, Parliamentarism, Multicameralism, and Multipartyism." British Journal of Political Science 25(3): 289–325.

*Data Sources*

Brolo, Javier. 2012. Resultados electorales de los partidos políticos en Guatemala de 1985 a 2012. Guatemala City: ASIES.

Brolo, Javier. 2013. El tamaño importa: Variaciones por distrito en elecciones legislativas de Guatemala (1984–2011). Guatemala City: ASIES.

Congreso de la República de Guatemala. n.d. “Decretos Legislativos.”
https://www.congreso.gob.gt/seccion_informacion_legislativa/decretos.

Secretaría General de la Presidencia de la República de Guatemala. 2025. Listado de vetos presidenciales compilado por la Secretaría General, 9 de octubre. Guatemala City: Gobierno de Guatemala.
Tribunal Supremo Electoral (TSE). n.d. “Memorias Electorales.”
https://www.tse.org.gt/images/uaip/memorias.electorales.php.

V-Dem Country-Year+Others Dataset
Coppedge, Michael, John Gerring, Carl Henrik Knutsen, Staffan I. Lindberg, Jan Teorell, David Altman, Fabio Angiolillo, Michael Bernhard, Agnes Cornell, M. Steven Fish, Linnea Fox, Lisa Gastaldi, Haakon Gjerløw, Adam Glynn, Ana Good God, Allen Hicken, Katrin Kinzelbach, Joshua Krusell, Kyle L. Marquardt, Kelly McMann, Valeriya Mechkova, Juraj Medzihorsky, Anja Neundorf, Pamela Paxton, Daniel Pemstein, Josefine Pernes, Johannes von Römer, Brigitte Seim, Rachel Sigman, Svend-Erik Skaaning, Jeffrey Staton, Aksel Sundström, Marcus Tannenberg, Eitan Tzelgov, Yi-ting Wang, Tore Wig, Steven Wilson and Daniel Ziblatt. 2026. "V-Dem Country-Year Dataset v16" Varieties of Democracy (V-Dem) Project. https://doi.org/10.23696/vdemds26.

V-Party Dataset
Lindberg, Staffan I., Nils Düpont, Masaaki Higashijima, Yaman Berker Kavasoglu, Kyle L. Marquardt, Michael Bernhard, Holger Döring, Allen Hicken, Melis Laebens, Juraj Medzihorsky, Anja Neundorf, Ora John Reuter, Saskia Ruth–Lovell, Keith R. Weghorst, Nina Wiesehomeier, Joseph Wright, Nazifa Alizada, Paul Bederke, Lisa Gastaldi, Sandra Grahn, Garry Hindle, Nina Ilchenko, Johannes von Römer, Steven Wilson, Daniel Pemstein, Brigitte Seim. 2022. Varieties of Party Identity and Organization (V–Party) Dataset V2. Varieties of Democracy (V–Dem) Project. https://www.v-dem.net/data/v-party-dataset/